### 2020-05-16 writeup/proof_transfer_ex.tex
proof

\begin{lemma}\label{lem_hat_v}
Suppose the assumptions of Proposition \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order. Then we have that the optimal ratio for $\val(w)$ satisfies
	$$|\hat w -1|=\OO(p^{-1}).$$%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{lemma}
\begin{proof}
The proof is also similar to the one for Proposition \ref{thm_cov_shift}. First it is easy to observe that $\val(w)\le \val(-w)$ for $w\ge 0$. Hence it suffices to consider the $w\ge 0$ case.

@@

@@

Recall that under the setting of Example \ref{ex_basic}, \eqref{Deltavar} and \eqref{Deltabeta} hold.  Then we get that $\te(\hat{\beta}_t^{\MTL})< \te(\hat{\beta}_t^{\STL})$ whp if
\be\label{pos1}\left(pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} + \bigo{p^{1/2+\e}d^2} \right)\cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 <  \sigma^2 \bigbrace{1 + \bigo{p^{-1/2 + \varepsilon}}};\ee
otherwise if
\be\label{neg1}\left(pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} + \bigo{p^{1/2+\e}d^2} \right)\cdot \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 >  \sigma^2 \bigbrace{1 + \bigo{p^{-1/2 + \varepsilon}}},\ee
then we have $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ whp.

>>>>

\begin{lemma}\label{lem_hat_v}
Suppose the assumptions of Proposition \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order. Then we have that the optimal ratio for $\val(w)$ satisfies
	\be\label{hatw_add1}|\hat w -1|=\OO(p^{-1}).\ee
	%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{lemma}
\begin{proof}
The proof is also similar to the one for Proposition \ref{thm_cov_shift}. First it is easy to observe that $\val(w)\le \val(-w)$ for $w\ge 0$. Hence it suffices to consider the $w\ge 0$ case.

@@

\begin{proof}[Proof of Proposition \ref{prop_var_transition}]

Under the setting of Proposition \ref{prop_var_transition}, we have
\begin{align*}
	\te(\hat{\beta}_t^{\MTL}) &=~ \hat{v}^2 \bignorm{ (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v} \beta_t)}^2 \nonumber \\
			&~ + \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } + (\sigma_1^2 -\sigma_2^2)\hat v^2  \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_1^\top X_1}.
\end{align*}
Using concentration of random vector with i.i.d. Gaussian entries, Lemma \ref{largedeviation}, we obtain that
\begin{align*}
	\te(\hat{\beta}_t^{\MTL}) &=~ \hat{v}^2 \left[d^2 +\left( w-1\right)^2\kappa^2\right]\bigtr{ (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2} \cdot \left(1+\OO(p^{-1/2})\right)\nonumber \\
	& + (\sigma_1^2 -\sigma_2^2)\hat v^2  \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_1^\top X_1} + \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }.
\end{align*}
In the setting of Proposition \ref{prop_var_transition}, we can also obtain that
$$|\hat w-1|=\OO(p^{-1}).$$
as in \eqref{hatw_add1}. We omit the details of the proof, since it is almost the same as the one in the proof of Lemma \ref{lem_hat_v}. Thus $\te(\hat{\beta}_t^{\MTL}) $ can be reduced to 
$$	\te(\hat{\beta}_t^{\MTL}) = \left[\Delta_\beta + \Delta_{diff}+  \sigma_2^2 \cdot \tr(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \right]\cdot \left(1+\OO(p^{-1/2})\right), $$
where $\Delta_\beta:= \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }$ has been estimated in \eqref{Deltabeta} and $\Delta_{diff}$ is
\begin{align*}	
	\Delta_{diff}:=(\sigma_1^2 -\sigma_2^2) \cdot \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_1^\top X_1} .
\end{align*}
Using \eqref{eq_isometric} and Lemma \ref{lem_cov_derivative}, a similar bound holds for $\Delta_{diff}$ as in \eqref{Deltabeta}:
$$\bigbrace{1 - \sqrt{\frac{1}{c_1}}}^2\le \frac{\Delta_{diff}}{\sigma_1^2 -\sigma_2^2}\cdot \left( \frac{c_1 (c_1+c_2)}{(c_1+c_2-1)^3} + \OO(p^{-1/2+\e})\right)^{-1}\le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^2.$$

Next we compare $\Delta_{\var}$ with $\Delta_\beta + \Delta_{diff}$. Under the condition $d^2 < \frac {2\sigma^2} {3p} \frac{(c_1 + c_2 -1)^2}{c_1 (c_1 + c_2)(c_2 - 1)}$, if $\sigma_1^2\le \sigma_2^2$ then we always have that $\Delta_{\vari} > \Delta_{\beta}$ and hence $\te(\hat{\beta}_t^{\MTL})$ is smaller than $\te(\hat{\beta}_t^{\STL})$. It remains to consider the case $\sigma_1^2 \ge \sigma_2^2$. If 
\begin{align} 
&\sigma_2^2  \frac{c_1}{(c_2-1)(c_1 + c_2 -1)}\left(1 + \bigo{p^{-1/2 + \varepsilon}}\right) \label{cond sigma1}\\
&> \left[\frac{c_1^2 (c_1+c_2)}{(c_1+c_2-1)^3} + \bigo{p^{-1/2 + \varepsilon}}\right]\left[pd^2 \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 + \frac{\sigma_1^2-\sigma_2^2}{c_1} \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^2\right] ,\nonumber
\end{align}
then we have $\Delta_{\vari} > \Delta_{\beta}$ and we have $\te(\hat{\beta}_t^{\MTL})$ is smaller than $\te(\hat{\beta}_t^{\STL})$. For $c_1\ge 100$, we see that \eqref{cond sigma1} is implied by the following condition
$$\sigma_1^2 \le pd^2 \cdot c_1 + \left[ 1+ \frac23\frac{(c_1 + c_2 - 1)^2}{(c_1 + c_2) (c_2 - 1)} \right]\sigma_2^2$$
as long as $p$ is large enough. On the other hand, if 
\begin{align} 
&\sigma_2^2  \frac{c_1}{(c_2-1)(c_1 + c_2 -1)}\left(1 + \bigo{p^{-1/2 + \varepsilon}}\right) \label{cond sigma12}\\
&< \left[\frac{c_1^2 (c_1+c_2)}{(c_1+c_2-1)^3} + \bigo{p^{-1/2 + \varepsilon}}\right]\left[pd^2 \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 + \frac{\sigma_1^2-\sigma_2^2}{c_1} \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^2\right] ,\nonumber
\end{align}
then we have $\Delta_{\vari} < \Delta_{\beta}$ and we have $\te(\hat{\beta}_t^{\MTL})$ is larger than $\te(\hat{\beta}_t^{\STL})$. For $c_1\ge 100$, we see that \eqref{cond sigma12} is implied by the following condition
$$\sigma_1^2 \ge  pd^2 \cdot c_1 + \left[ 1+ \frac32\frac{(c_1 + c_2 - 1)^2}{(c_1 + c_2) (c_2 - 1)} \right]\sigma_2^2.$$
\end{proof}

@@

Recall that under the setting of Example \ref{ex_basic}, \eqref{Deltavar} and \eqref{Deltabeta} hold.  Then we get that $\te(\hat{\beta}_t^{\MTL})< \te(\hat{\beta}_t^{\STL})$ whp if
\be\label{pos1}\left(pd^2 \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} + \bigo{p^{1/2+\e}d^2} \right)\bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 <  \sigma^2 \bigbrace{1 + \bigo{p^{-1/2 + \varepsilon}}};\ee
otherwise if
\be\label{neg1}\left(pd^2 \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} + \bigo{p^{1/2+\e}d^2} \right)\bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 >  \sigma^2 \bigbrace{1 + \bigo{p^{-1/2 + \varepsilon}}},\ee
then we have $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ whp.

### 2020-05-16 writeup/transfer.tex
proof

Specifically, the transition threshold is derived in the following proposition.
\begin{proposition}\label{prop_dist_transition}
	In the setting of Example \ref{ex_basic} with $\sigma_1 = \sigma_2 = \sigma$, assume that $c > 100$ is a fixed constant.
	Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between the model distance and the noise level:
	\begin{itemize}
		\item If ${d^2} < \frac {2\sigma^2} {3p} \frac{(c_1 + c_2 -1)^2}{c_1 (c_1 + c_2)(c_2 - 1)} - o(1)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If ${d^2} \ge \frac {3\sigma^2} {2p} \frac{(c_1 + c_2 -1)^2}{c_1(c_1 + c_2)(c_2 - 1)} + o(1)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}

@@

\begin{proposition}\label{prop_var_transition}
	In the setting of Example \ref{ex_basic} with $d$ being fixed but $\sigma_1$ varies, we derive the following transition as a parameter of $\sigma_1$:
	\begin{itemize}
		\item If $\sigma_1^2 \le p d^2 \cdot c_1 + \frac{(c_1 + c_2 - 1)^2}{(c_1 + c_2) (c_2 - 1)}\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\sigma_1^2 > p d^2 \cdot c_1 + \frac{(c_1 + c_2 - 1)^2}{(c_1 + c_2) (c_2 - 1)} \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}

>>>>

Specifically, the transition threshold is derived in the following proposition.
\begin{proposition}\label{prop_dist_transition}
	In the setting of Example \ref{ex_basic} with $\sigma_1 = \sigma_2 = \sigma$, assume that $c_1 > 100$ is a fixed constant.
	Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between the model distance and the noise level:
	\begin{itemize}
		\item If ${d^2} < \frac {2\sigma^2} {3p} \frac{(c_1 + c_2 -1)^2}{c_1 (c_1 + c_2)(c_2 - 1)}$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If ${d^2} \ge \frac {3\sigma^2} {2p} \frac{(c_1 + c_2 -1)^2}{c_1(c_1 + c_2)(c_2 - 1)} $, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}

@@

\begin{proposition}\label{prop_var_transition}
	In the setting of Example \ref{ex_basic} with $d$ being fixed but $\sigma_1$ varies, assume that $c_1 > 100$ is a fixed constant and $d^2 < \frac {2\sigma^2} {3p} \frac{(c_1 + c_2 -1)^2}{c_1 (c_1 + c_2)(c_2 - 1)}$. Then we derive the following transition as a parameter of $\sigma_1$:
	\begin{itemize}
		\item If $\sigma_1^2 \le p d^2 \cdot c_1 +\left[1+ \frac23\frac{(c_1 + c_2 - 1)^2}{(c_1 + c_2) (c_2 - 1)}\right]\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\sigma_1^2 > p d^2 \cdot c_1 +\left[1+ \frac32\frac{(c_1 + c_2 - 1)^2}{(c_1 + c_2) (c_2 - 1)}\right] \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}
